---
title: Tools to Support Relative Importance Analysis
output: github_document
---

[![Stable version](http://www.r-pkg.org/badges/version-last-release/domir)](https://cran.r-project.org/package=domir) [![downloads](http://cranlogs.r-pkg.org/badges/grand-total/domir)](https://cran.r-project.org/package=domir)

```{r setup, include = FALSE}
devtools::load_all()
```

# Overview

The `{domir}` package contains functions that apply decomposition-based relative importance analysis methods (*dominance analysis* or *Shapley value decomposition*) to predictive modeling functions in R.

The intention of this package is to provide a flexible user interface to dominance analysis—a popular relative importance analysis method that extends on the rigorous solution concept of Shapley value decomposition in cooperative game theory. The user interface is structured such that `{domir}` automates the decomposition of the returned value and comparisons between model inputs and the user provides the model inputs, the predictive model into which they are entered, and returned value from the model to decompose.

# Installation

To install the most recent stable version of `{domir}` from CRAN use:

`install.packages("domir")`

To install the working, development version of `{domir}` using the `{devtools}` package use:

`devtools::install_github("https://github.com/jluchman/domir")`

Coming soon: see the `{domir}`-based `dominance_analysis` function for the [{parameters}](https://github.com/easystats/parameters) package from the `{easystats}`/easyverse.

# What `{domir}` Does

The primary dominance analysis function `domir` implements the most computationally intensive and programming heavy parts of dominance analysis for the user and has relatively few requirements on the predictive modeling functions with which it can work.

The flexibility of `domir` comes at the cost of more complexity for the user in terms of setting up a function that accepts the type of input `domir` will provide (currently only a 'formula') and and expects to receive (currently only a numeric scalar).

Below these ideas are outlined in greater detail in the context of a few examples. The next section begins the discussion with a more extensive comparison of `domir` with packages that implement similar methods.

## Comparison with Existing Relative Importance Packages

The `domir` function is similar to the "lmg" type for the `calc.relimpo` function in the `{relaimpo}` package as well as the `dominanceAnalysis` function in the [`{dominanceanalysis}` package](https://github.com/clbustos/dominanceAnalysis) (not on CRAN). `domir` can replicate the results produced by both the above packages but, as will be seen, requires more user input.

To illustrate these points, consider the following example linear regression on which all three of the dominance analysis results to come are based:

`lm(mpg ~ am + vs + cyl, data = mtcars)`

Classic dominance analysis uses the variance explained $R^2$ as fit statistic (i.e., as implemented by `lm`'s `summary` method) and so will this example.

### `{domir}`'s `domir`

Implementing a 'classic' dominance analysis on this linear regression in `domir` can be inputted as:

```{r init_domin}
lm_wrapper <-       
  function(formula, data) {
    result <- 
      lm(formula, data = data) |> 
      summary()
    return(result[["r.squared"]])
  }

domir(mpg ~ am + vs + cyl, 
      lm_wrapper,
      data = mtcars)
```

In `domir`, the `lm` model is not submitted directly. Rather, it is wrapped into a function (i.e., `lm_wrapper`) that, in this case, accepts two arguments; *formula* or an R formula and *data* a data frame in which the independent variables in the formula are present. The result of the `lm` is piped (i.e., `|>`) into the `summary` function and the result is captured as the object *result*. The *result* object is then filtered to just the **r.squared** element and returned.

What `domir` does automate taking subsets of the *formula* and submit them, repeatedly until all possible subsets have been submitted, to `lm_wrapper` (see this [vignette](https://CRAN.R-project.org/package=domir/vignettes/domir_basics.html) for a conceptual discussion of dominance analysis). In this way, `domir` is a `Map`- or `lapply`-like function as it receives an object on which to operate (i.e., the *formula*) and a function to which to apply to it. `domir` expects a numeric scalar to be returned from the function.

Like `lapply`, other arguments (`data = mtcars`) can also be passed to each call of the function and must be explicitly built into the wrapper function.

What is important to note about `domir` that differs from other dominance analysis-oriented functions discussed below is that `domir` expects that the user will supply the analysis pipeline linking the *formula* it passes to the numeric scalar value that it expects. This 'supply the pipeline' approach makes `domir` far more flexible than other implementations but does require the user to think more carefully about how to structure the pipeline.

Note that the focus of `domir`'s `print`-ed results focuses on the numerical results from "General Dominance Values" and "Conditional Dominance Values" and, a logical matrix of "Complete Dominance Designations".

See also the (now superseded) `domir::domin` function for another approach to structuring the input pipeline for dominance analysis.

### `{relaimpo}`'s `calc.relimp` with `type = "lmg"`

`{relaimpo}` is not a dominance analysis software but does produce general dominance value decomposition for linear regression using the explained variance $R^2$ in the `calc.relimp` function with the argument `type = "lmg"`.

```{r init_relaimpo}
relaimpo::calc.relimp(mpg ~ am + vs + cyl, 
                      data = mtcars, 
                      type = "lmg")
```

`calc.relimp` has a similar to structure to that of `domir` but does not require a pipeline function. This is because `{relaimpo}` is specialized and works only with `lm` models and the variance explained $R^2$ as a fit statistic. `calc.relimp` also allows for multiple methods of submitting (i.e., correlation matrices, fitted `lm` object, a `data.frame`) given that it always implements the same model.

`calc.relimp`'s printed results provide relative importance metric values that match those obtained from `domir` (i.e., the general dominance values). In addition, `calc.relimp` reports the average `lm` coefficients across numbers of independent variables/$X$s in a way similar to the conditional dominance values reported by `domir`—an additional and useful result to show the impact of inclusion of different numbers of independent variables on obtained coefficients/predicted values.

Again, note that `{relaimpo}` is not dominance analysis-oriented and does not report on dominance designations or dominance values other than the general dominance values.

### `{dominanceanalysis}`'s `dominanceAnalysis`

`{dominanceanalysis}` implements dominance analysis for several different predictive models including `lm` , `betareg`, and `glm` each with its own built-in (pseudo-)$R^2$.

```{r init_da}
lm_model <- lm(mpg ~ am + vs + cyl, 
      data = mtcars)

dominanceanalysis::dominanceAnalysis(lm_model)
```

`dominanceAnalysis` has a simpler approach than `domir` to get a 'classic' dominance analysis as it accepts a fitted `lm` model as input and requires use of the explained variance $R^2$ as the returned value. The object returned by `dominanceAnalysis` is large and contains the fit statistic values from all subsets as well as computed dominance statistics based on them. Several helper functions are available to extract specific dominance and other results for printing to the console.

`dominanceAnalysis`'s default printed output is focused on qualitative dominance designations but also reports a sorted, average contribution metric (i.e., general dominance values).

As mentioned above, `{dominanceanalysis}` can be used with around seven different predictive models and implements a (pseudo-)$R^2$ as returned values for each. Itis worth noting that the upcoming `dominance_analysis` function in the `{parameters}` package takes a similar approach as `{dominanceanalysis}` but works from the `{insight}` engine linked with `performance::r2` which allows extension to many different models.

## How `{domir}` Extends on Previous Packages

The intention of `{domir}` is to extend relative importance analysis to new data analytic situations the user might encounter where a decomposition-based relative importance method such as dominance analysis could be valuable.

The sections below outline some pertinent examples that the `domir` function can accommodate that cannot be r

### Linear Model Revisited

Given that the user supplies the analysis pipeline, one component of `domir`'s flexibility is in allowing the user to apply any applicable fit statistic as a returned value for the purposes of relative importance analysis.

In the example below, the explained variance $R^2$ is swapped with an alternative, but nonetheless applicable, fit statistic: the McFadden pseudo-$R^2$ as implemented by the `{pscl}` package.

The example below is more complex than the previous `domir` call as the analysis pipeline is submitted as an anonymous function with a single argument (*fml*). In part, this approach is taken to show that the user *can* submit the function to `domir` in this way. In addition, note that the `data` argument submitted to the `lm` function is built-into the analysis pipeline instead of passed as an argument; both are valid methods of setting arguments to predictive analyses.

```{r lm_w_mcf}
domir(mpg ~ am + vs + cyl, 
      \(fml)
      {(result <- 
          lm(fml, data = mtcars) |> 
          pscl::pR2()
      ) |> capture.output()
        return(result[["McFadden"]])
      })
```

The use of the McFadden pseudo-$R^2$ has produced effectively the same answers, in terms of qualitative importance inferences about the independent variables, as that of the dominance analysis using the explained variance $R^2$.

It is also worth noting that the use `capture.output` in the anonymous function was not not strictly necessary. If not used, `domir` will print far more output than is needed as `pscl::pR2` is a rather verbose function and will print a message for each model fitted.

### Ordered Logistic Regression

The user-defined analysis pipeline also allows for extending predictive modeling to effectively any predictive model (that the user can adapt the formula input to accommodate). The example below is applied to the `polr` function from the `{MASS}` package using `peformance::r2`'s result as a returned value.

```{r da_polr}
mtcars2 <- data.frame(mtcars, carb2 = as.factor(mtcars$carb))

domir(carb2 ~ am + vs + mpg, 
      \(fml) 
      MASS::polr(fml, data = mtcars2) |>
        performance::r2() |> unlist()
) 
```

The call to `unlist` in the anonymous function above ensures that the returned value is a numeric scalar as opposed to a list with a single element.

### Random Forest

`domir` can also work with predictive models that do not produce model coefficients like `randomForest::randomForest`. The dominance analysis approach's results differ from the built-in variable importance method plotted below (which is arguably better suited for model selection) but can, and in the case of many of the variables do, agree on which independent variables are important.

The dominance analysis here is based on a squared correlation of the predicted values with the dependent variable (i.e., the explained variance $R^2$).

```{r da_rpart}
set.seed(5621)

rf_model <- 
  randomForest::randomForest(mpg ~ am + qsec + cyl, data = mtcars, 
                             importance = TRUE)

data.frame(`%IncMSE` = rf_model$importance[,1], `RankIncMSE` = rank(rf_model$importance[,1]*-1), `IncNodePurity` = rf_model$importance[,2], `RankIncPurity` = rank(rf_model$importance[,2]*-1),check.names = FALSE)

cor(predict(rf_model), mtcars$mpg)^2

domir(mpg ~ am + qsec + cyl, 
      \(fml) {
        set.seed(5621)
        result <- 
          randomForest::randomForest(fml, data = mtcars, 
                             importance = TRUE)
        cor <- cor(predict(result), mtcars$mpg)
        return(cor^2)
      }
)
```

Note the use of `set.seed` prior to all calls to `randomForest`. These ensure that the random processes within the `randomForest` function result in a reproducible set of predicted values (and $R^2$ metric). The calls to individual `randomForest`s also had to use the `importance = TRUE` argument (though they are not used) to ensure matching with the original result as they affect the state of the random number generator.

### Zero-Inflated Poisson

One distinct advantage of having the level of flexibility in the analytic pipeline that `domir` offers is that this that it can work directly with modeling functions that are more complex. The example below uses the `zeroinfl` model from the package `{pscl}` that accepts a `Formula::Formula` object (i.e., a multi-equation formula) instead of a standard R formula.

The below example uses the entries in the formula to plug into the `Formula` object that will be submitted to the `zeroinfl` model.

```{r da_zip}
library(Formula)

domir(~ fem + mar + kid5, 
      \(fml) {
        result <- 
          as.Formula(fml, fml) |> 
          update(art ~ .) |>
          pscl::zeroinfl(data = pscl::bioChemists) |>
          performance::r2()
        return(result[["R2"]])
      })
```

In this example, note the absence of a dependent variable in the model formula. `domir` does not require a left hand side/dependent variable to accommodate situations like the one here where it is added later in the analysis pipeline. Also, note that the *fml* passed to the pipeline is repeated in the Poisson and inflation model equations and then adapted to a `Formula` object before submitting to `zeroinfl`.

# Further Reading

*...to be populated...*
