---
title: Tools to Support Relative Importance Analysis
output: github_document
---

```{r setup, echo=FALSE, message=FALSE}
devtools::load_all()
```

# Overview

The intention of the `{domir}` package is to provide tools that allow relative 
importance analysis across a wide variety of practical data analytic situations.
With `{domir}`, if you have a statistical/machine learning model and an extractor 
function to obtain a fit metric, you can conduct a relative importance analysis.

More specifically, `{domir}` is intended to implement a set of flexible wrapper 
and helper functions for conducting relative importance analysis and the 
current implementation of the package has a focus on dominance analysis with 
the `domin` function as a relative importance analysis method.  Currently, 
`domin` is the only importance function that is implemented.

For readers looking to familiarize themselves more with the dominance 
analysis methodology, a more extensive conceptual discussion of dominance 
analysis (which focuses on the Stata version of `domin`) as a method is 
available [here](https://github.com/jluchman/domin/blob/master/README.md).

# Installation

To install the most recent stable version of `domir` from CRAN use:

`install.packages("domir")`

To install the working development version of `{domir}` using the `devtools` 
package use:

`devtools::install_github("https://github.com/jluchman/domir")`

# What `{domir}` Does

Before discussing details of the `{domir}` package, I provide some examples of 
what `{domir}` can do.

The focus of this section is on outlining how `domir::domin` extends existing 
packages and on the structure of the function.

## Comparison with Existing Relative Importance Packages

Fundamentally, `domir::domin` is an extension of the "lmg" type for the 
`calc.relimpo` function in the `{relaimpo}` package as well as the 
`dominanceAnalysis` function in the `{dominanceanalysis}` package.

`domir::domin` can replicate the results produced by the above packages but, as 
will be seen, requires a "deconstructed" the model to be submitted to it.  This 
difference in structure does make `domin` more complex but also allows the 
function a great deal more flexibility in terms of the kinds of models and 
fit metrics that can be dominance analyzed.

Before discussing some of the elements that make `domin` flexible, consider the 
following example that shows how `domin` is similar to existing packages. All 
three of the dominance analysis results to come are based on the following 
linear model:

`lm(mpg ~ am + vs + cyl, data = mtcars)`

The variance explained $R^2$ will be the focal fit metric.

### `{domir}`'s `domin`

```{r init_domin}
domin(mpg ~ am + vs + cyl, 
      lm, 
      list(summary, "r.squared"), 
      data = mtcars)
```

In `domin`, the `lm` model is submitted in pieces.  Specifically, the key inputs 
were the formula (`mpg ~ am + vs + cyl`) and the model to be called using 
the formula (`lm`).  In this way, `domin` is a `Map`- or `apply`-like 
function as it receives an object on which to operate (i.e., the formula) and 
a function to which to apply to it.

In being like `apply`, `domin` is agnostic to the fit metric to use for the 
model called and it must be supplied with a list outlining extractor function 
information (`list(summary, "r.squared")`; described further in the
[Details](#Details) section).

Like `apply`, other arguments (`data = mtcars`) can also be passed to each 
call of `lm`.

The focus of `domin`'s `print`-ed results focuses on the numerical results from 
"General Dominance Statistics" and "Conditional Dominance Statistics" and, a 
logical matrix of "Complete Dominance Designations".

### `{relaimpo}`'s `calc.relimp` with "lmg"

``` {r init_relaimpo}
relaimpo::calc.relimp(mpg ~ am + vs + cyl, 
                      data = mtcars, 
                      type = "lmg")
```

`{relaimpo}`'s `calc.relimp` accepts only `lm` models and the variance explained 
$R^2$ as a fit metric.  As a result, the function does not ask for model
or fit metric type.

The function's printed results provide the "lmg" relative importance statistics 
(i.e., General Dominance Statistics) and, in addition, report the average `lm` 
coefficients across all models.

### `{dominanceanalysis}`'s `dominanceAnalysis`

``` {r init_da}
dominanceanalysis::dominanceAnalysis(lm(mpg ~ am + vs + cyl, 
                                        data = mtcars))
```

`{dominanceanalysis}`'s `dominanceAnalysis` implements dominance analysis 
for specific models, of which `lm` is a supported model. 
`dominanceAnalysis` accepts a fitted `lm` object as input and uses the 
explained variance $R^2$ as the fit metric.

`dominanceAnalysis`'s printed output is focused on qualitative dominance 
designations but also reports the, magnitude sorted, average contribution
(i.e., General Dominance Statistic) values.

## How `{domir}` Extends on Previous Packages

The intention of `{domir}` is to extend relative importance to new data 
analytic situations the user might encounter where a dominance analysis 
could be valuable.

The sections below outline some pertinent examples of specific models that the
`domin` function can accommodate.

### Linear Model Revisited

`domin` is fit metric agnostic and, as such, one component of its flexibility 
is in allowing the user to apply any applicable fit metric for a model for
the purposes of relative importance analysis.

In this example, the explained variance $R^2$ is swapped with an alternative, 
but nonetheless applicable, fit metric: the McFadden pseudo-$R^2$ as 
implemented by the `{pscl}` package.

Note the use of the pipes to `capture.output` and `invisible`.  These are not
not strictly necessary but if not used will print far more output than is 
needed as `pscl::pR2` is a rather verbose function and will print a message for 
each model fitted.

```{r lm_w_mcf}
(mcf_da_lm <- 
   domin(mpg ~ am + vs + cyl, 
         lm, 
         list(pscl::pR2, "McFadden"), 
         data = mtcars)) |> 
  capture.output() |> 
  invisible()

mcf_da_lm
```

Note that this fit metric produces effectively the same answers, in terms of 
qualitative importance inferences about the terms, as that from the 
explained variance $R^2$.

### Ordered Logistic Regression

`domin` acts like an `apply` function for models and does not have built in 
methods. This is another component of its flexibility as it can  accommodate 
functions that, to this point, have not been supported in relative importance 
analysis.  One pertinent example is the `polr` function from the `{MASS}` 
package also using `pscl::pR2` as fit metric.

```{r da_polr}
mtcars2 <- data.frame(mtcars, carb2 = as.factor(mtcars$carb))

(da_polr <- 
    domin(carb2 ~ am + vs + mpg, 
          MASS::polr, 
          list(pscl::pR2, "McFadden"), 
          data = mtcars2)) |> 
  capture.output() |> 
  invisible()

da_polr
```

### Decision Trees

`domin` can also accept models that do not produce model coefficients like 
`rpart::rpart`. 

```{r da_rpart}
domin(mpg ~ am + vs + cyl, 
      rpart::rpart, 
      list(\(model) 
            list(R2 = 1-model$cptable[nrow(model$cptable), 3]), 
            "R2"),
      data = mtcars)
```

Note that an anonymous function can be used as a valid submission to 
the `fitstat` argument.  In this case, the anonymous function transforms and 
extracts the proportion of error from the `rpart` object.  If the model 
object returns its own fit statistic, it can be extracted using an 
anonymous function.

### Multinomial Logistic (softmax) Regression with Extra Features

`domin`, similar to other packages, can combine multiple terms into a 
single set as well as use one or more terms as covariate(s) in all model 
subsets.

This example outlines another model, `multinom` from the `{nnet}` package,  
another function that has not been accommodated in relative importance 
packages, that uses sets and all/covariate terms.

In addition, `complete = FALSE` which saves a little computation time and 
suppresses reporting complete dominance designations.

```{r da_multinom}
(da_mnl <- 
  domin(carb2 ~ mpg, 
      nnet::multinom, 
      list(pscl::pR2, "McFadden"), 
      sets = list(c("am", "vs"), c("cyl", "disp")),
      all = c("gear"),
      complete = FALSE,
      data = mtcars2)) |> 
  capture.output() |> 
  invisible()

da_mnl

da_mnl$Subset_Details$Full_Model
```

The `domin` automatically combines the entries in the `formula_overall`, 
`sets`, and `all` arguments. The full model formula can be obtained from the 
`domin` object in the `.$Subset_Details$Full_Model` element.

### Zero-Inflated Poisson with Wrapper Function

Although `domin` can work directly with modeling functions that accept 
standard formula, more complex formulas such as those used by models such as 
`zeroinfl` models from the package `{pscl}` can also be accommodated using 
wrapper functions.

The below wrapper function`zinfl_wrap` uses the entries in the formula to 
create a symmetric count and zero-inflation formulas that will be submitted 
to `zeroinfl` model.  

In an effort to illustrate what each model submitted to `zeroinfl` looks like,
the model formula for all 7 models is printed before each run.

```{r da_zip}
zinfl_wrap <- function(model, ...) {
  zip_terms <- model |> terms() |> attr("term.labels") |> paste(collapse = " + ")
  zip_formula_rhs <- zip_terms |> rep(times = 2) |> paste(collapse = " | ")
  zip_formula_lhs <- (model |> all.vars())[[1]]
  zip_formula <- c(zip_formula_lhs, zip_formula_rhs) |> paste(collapse = " ~ ") |> as.formula()
  print(deparse(zip_formula))
  pscl::zeroinfl(zip_formula, ...)
}

domin(art ~ fem + mar + kid5, 
      zinfl_wrap,
      list(\(model) {capture.output(result <- pscl::pR2(model)); result}, "McFadden"), 
      data=pscl::bioChemists)
```

Further discussion of how to generate wrapper commands is outlined below in the 
[Details](#Details) section.

----

# Details

Having provided some details of what `domin` is capable of, this section moves 
on to outline details of how `domin` works by way of the structure of the 
function.

As is noted above, `domin` takes inspiration from the `apply` family of 
functions and works in a similar way, creating the dominance analysis from 
three 'building block' arguments:

1. a formula (`formula_overall`)
2. a modeling function (`reg`)
3. list of instructions to call an extractor function that obtains a model fit metric (`fitstat`)

The three arguments above effectively invoke the following process.

`reg(formula_overall) |> fitstat()`

Hence, the `reg` function is called using the `formula_overall` as argument and 
the results of the `reg` model are are 'piped' to `fitstat` to obtain the 
fit statitsic that is used for dominance statistic computation.

In the sections below, each argument to `domin` is discussed in greater detail.

## Formula

The first, and most important, argument for `domin` is the formula inputs and 
understanding how it is parsed is important for the effective use of `domin`.

The formula inputs are derived from three places in `domin`.

a. the `formula_overall` argument
b. the `sets` argument
c. the `all` argument

*** section in progress ***

### `formula_overall`

The `formula_overall` argument must take the form of `response ~ terms` as 
that is a standard format for many modeling functions such as `lm` and `glm`. 

The entries on the right hand side of the formula are parsed using the `terms` 
utilities in the `stats` library. The desired behavior is that variable names 
separated by `+` are divided up and used as terms in computing all combinations.

It is also important to note that all special formula processing is applied 
to the formula including the use of `I()`, `:`, `*`, and `offset()`.  
The actual list of terms on which `domin` computes all combinations is 
obtained from `attr(,"term.labels")`.  If the user wants to test to see what 
`domin` will do with the formula submitted use: 
`formula(.) |> terms() |> attr("term.labels")` to see which labels it will 
produce (note also that `domin` does not have a method to accommodate second 
or higher order terms like `{relaimpo}` and will issue a warning when 
second or higher order terms are used).

### `sets`

After processing, the formula object is combined with entries from the `sets` 
argument.  The entries in the `sets` argument are minimally processed and 
are combined



and is not 'data frame aware'.  That is, shorthand such as `~ .` 
will not work to select variables in a data frame even if a `data` argument 
is supplied to the `domin` function.

```r{formula_example}
formula(mpg ~ .) |> terms(data=mtcars) |> formula()
```


...note all and sets go here as well

many modeling functions that accept a formula that follow the 
standard `response ~ terms` format. 

`domin` works closely with the `do.call` function and the structure of the 
`domin` function is similar as a result.  Users are encouraged to 
read the documentation for the `do.call` function to understand how `domin` is 
implemented and for bug/error checking.

wrapper function that can be used with 
The format used by domin can be extended 
to other functions focused on independent variable/predictor/feature-based RI 
can be accommodated with a additional wrapper functions based 
on the formula it creates and submits to modeling functions.  Some examples are
offered below.


## Modeling Function

## Fit Statistic Extractor Function

# Consideraions

This section outlines a few key considerations 

`domin` does _not_ check to ensure that the sample underlying the modeling 
is consistent across modeling runs on which the dominance statistics are 
computed.  If the modeling function uses a `na.action` that omits missing 
responses, and different variables have different missing observations, the 
sample included for each modeling run will vary. I recommend filtering the 
data to include _only_ the sample that does not have any missing data on the 
variables included.

In my view, `domin` is a decision tool for model explanation (i.e., 
understanding a fit, selected model) and is not as useful for model selection 
or inference.  Hence, if the user
