---
title: Tools to Support Relative Importance Analysis
output: github_document
---

```{r setup, echo=FALSE, message=FALSE}
devtools::load_all()
```

# Overview

The intention of the `{domir}` package is to provide tools that allow relative 
importance analysis across a wide variety of practical data analytic situations.
With `{domir}`, if you have a statistical/machine learning model and an extractor 
function to obtain a fit metric, you can conduct a relative importance analysis.

More specifically, `{domir}` is intended to implement a set of flexible wrapper 
and helper functions for conducting relative importance analysis and the 
current implementation of the package has a focus on dominance analysis with 
the `domin` function as a relative importance analysis method.

For readers looking to familiarize themselves more with the dominance 
analysis methodology, a more extensive conceptual discussion of dominance 
analysis (which focuses on the Stata version of `domin`) as a method is 
available [here](https://github.com/jluchman/domin/blob/master/README.md).

# Installation

To install the most recent stable version of `domir` from CRAN use:

`install.packages("domir")`

To install the working development version of `{domir}` using the `devtools` 
package use:

`devtools::install_github("https://github.com/jluchman/domir")`

# What `{domir}` Does

Before discussing details of the `{domir}` package, I provide some examples of 
what `{domir}` can do.

The focus of this section is on outlining how `domir::domin` extends existing 
packages and on the structure of the function.

## Comparison with Existing Relative Importance Packages

Fundamentally, `domir::domin` is an extension of the "lmg" type for the 
`calc.relimpo` function in the `{relaimpo}` package as well as the 
`dominanceAnalysis` function in the `{dominanceanalysis}` package.

`domir::domin` can provide the same results but is structured differently 
and, as will be discussed, is designed for flexibility - which does come 
at the expense of complexity.

All three of the models to come are based on the following linear model:

`lm(mpg ~ am + vs + cyl, data = mtcars)`

### `{domir}`'s `domin`

```{r init_domin}
domin(mpg ~ am + vs + cyl, 
      lm, 
      list(summary, "r.squared"), 
      data = mtcars)
```

Note that the `lm` model is submitted in pieces.  Specifically, the key inputs 
were the formula (`mpg ~ am + vs + cyl`), the indication of the model to be 
called (`lm`).  I recommend thinking of `domin` as a `Map`- or `apply`-like 
function given this structure.

A key difference as you will see from other similar functions is that `domin` 
must be supplied with a list outlining extractor function 
information (i.e., `list(summary, "r.squared")`; described further in the
[Details](#Details) section).

Like `apply`, other arguments (i.e., `data = mtcars`) can also be passed to each 
call of `lm`.

The focus of `domin`'s `print`-ed results focuses on the numerical results from 
"General Dominance Statistics" and "Conditional Dominance Statistics" and, 
optionally, a logical matrix of "Complete Dominance Designations".

### `{relaimpo}`'s `calc.relimp`

``` {r init_relaimpo}
relaimpo::calc.relimp(mpg ~ am + vs + cyl, 
                      data = mtcars, 
                      type = "lmg")
```

`{relaimpo}`'s `calc.relimp` is similar in structure to `domin` but only allows 
for `lm`s as well as the standard variance explained $R^2$ value.

The function also focuses only on the analogous statistics to 
"General Dominance Statistics".  In addition, it supplies, by default, average 
`lm` coefficients across all models.

### `{dominanceanalysis}`'s `dominanceAnalysis`

``` {r init_da}
dominanceanalysis::dominanceAnalysis(lm(mpg ~ am + vs + cyl, 
                         data = mtcars))
```

`{dominanceanalysis}`'s `dominanceAnalysis` accepts a fitted `lm` object (as 
does `calc.relimp`) and otherwise is similar in structure to both previous 
functions.  `dominanceAnalysis` also has built-in methods for specific model 
implementations like `lm` - which assumes the explained variance $R^2$ will be 
used.

`dominanceAnalysis`'s output is focused on qualitative dominance designations 
but also reports the "General Dominance Statistic" values.

## How `{domir}` Extends on Previous Packages

The intention of `{domir}` is to extend relative importance to new data 
analytic situations the user might encounter where a dominance analysis 
could be valuable.

The sections below outline some pertinent examples of specific models that the
`domin` function can accommodate.

### Linear Model Revisited

Because `domin` does not require the use of a specific fit metric, one 
important way in which it differs from the previous functions is in that 
it allows for different fit metrics than the standard error 
variance $R^2$ with models like `lm`.  In the example below, the 
the `lm` discussed above is dominance analyzed with the McFadden pseudo-$R^2$ 
as implemented by the `{pscl}` package.

Note the use of the `purrr::quietly` function to "mute" `pscl::pR2`.  This 
is not strictly necessary but if not used will print far more output than is 
needed as it is a rather verbose function.

```{r lm_w_mcf}
qui_pR2 <- function(model) purrr::quietly(pscl::pR2)(model)$result

domin(mpg ~ am + vs + cyl, 
      lm, 
      list(qui_pR2, "McFadden"), 
      data = mtcars)
```

The ability to use any fit metric that is desired is one part of the utility of 
the `domin` function.  

Note that this fit metric produces effectively the same answers, in terms of 
qualitative dominance inferences, as that from the explained variance $R^2$.

### Ordered Logistic Regression

Additionally, because `domin` does not have built in methods, it can 
accommodate functions that, to this point, have not been supported in 
relative importance analysis.  One pertient example is the `polr` function 
from the `MASS` package.

```{r da_polr}
mtcars2 <- data.frame(mtcars, carb2 = as.factor(mtcars$carb))

domin(carb2 ~ am + vs + mpg, 
      MASS::polr, 
      list(qui_pR2, "McFadden"), 
      data = mtcars2)
```

The ability to call any model with `domin` is another part of the function's 
utility.  

One particularly useful side-effect of relative importance analysis with 
non-linear models like `polr`is that the the log-odds transformed 
coefficients it produces are difficult to compare to one another whereas the 
dominance statistics produced by `domin` are much simpler to compare.

### Decision Trees

Given its flexibility, `domin` can also accept models that have 
traditionally not been dominance analyzed such as those in `{rpart}`.

```{r da_rpart}
domin(mpg ~ am + vs + cyl, 
      rpart::rpart, 
      list(\(model) 
            list(R2 = 1-model$cptable[nrow(model$cptable), 3]), 
            "R2"),
      data = mtcars)
```

It is worth noting are that the `fitstat` argument in the above 
`domin` call accepted an anonymous function (i.e., in R 4.1's `\()` 
notation) that effectively just extracted the proportion of error 
from the `rpart` object.  Hence, if the model object returns its own fit 
statistic, it can be extracted using an anonymous function.

### Multinomial Logistic (softmax) Regression with Extra Features

`multinom` from the `{nnet}` package is another good example model 
function that has not been accommodated in relative importance packages 
and is the model function featured in this example.

In this model, several sets of predictors are used as well as a predictor as 
covariate in all model subsets.

In addition, `complete = FALSE` which saves a little computation time and 
suppresses reporting complete dominance designations.

The `domin` automatically combines the entries in the `formula_overall`, 
`sets`, and `all` arguments. The full model formula looks like:

`carb2 ~ mpg + am + vs + cyl + disp + gear`

```{r da_multinom}
domin(carb2 ~ mpg, 
      nnet::multinom, 
      list(qui_pR2, "McFadden"), 
      sets = list(c("am", "vs"), c("cyl", "disp")),
      all = c("gear"),
      complete = FALSE,
      data = mtcars2, 
      trace = FALSE)
```

The ability to use sets of variables or covariates in all subsets is not a 
novel feature but nonetheless useful to illustrate as combined with `domin`'s 
other features.

### Zero-Inflated Poisson with Wrapper Function

Although `domin` can work directly with modeling functions that accept 
standard formula, more complex formulas such as those used by models such as 
`zeroinfl` models from the package `pscl` can also be accommodated using 
wrapper functions.

The below wrapper function`zinfl_wrap` uses the entries in the formula to 
create a symmetric count and zero-inflation formulas (with the new forward 
pipe operator in R 4.1) that will be submitted to `zeroinfl` model.  

In an effort to illustrate what each model submitted to `zeroinfl` looks like,
the model formula for all 9 models is printed before each run.

```{r da_zip}
zinfl_wrap <- function(model, ...) {
  zip_terms <- model |> terms() |> attr("term.labels") |> paste(collapse = " + ")
  zip_formula_rhs <- zip_terms |> rep(times = 2) |> paste(collapse = " | ")
  zip_formula_lhs <- (model |> all.vars())[[1]]
  zip_formula <- c(zip_formula_lhs, zip_formula_rhs) |> paste(collapse = " ~ ") |> as.formula()
  print(deparse(zip_formula))
  pscl::zeroinfl(zip_formula, ...)
}

domin(art ~ fem + mar + kid5, 
      zinfl_wrap,
      list(qui_pR2, "McFadden"), 
      data=pscl::bioChemists)
```

Further discussion of how to generate wrapper commands is outlined below in the 
[Details](#Details) section.

----

# Details

** section in progress **

`{domir}` implements dominance analysis using the `domin` function.

`domin` acts as a `Map`- or `apply`-like function that takes three 'building 
block' arguments:

1. a formula (`formula_overall`)
2. a modeling function (`reg`)
3. list of instructions to call an extractor function that obtains a model fit metric (`fitstat`)

The three arguments above effectively invoke the following process (using the 
new `|>` forward pipe operator notation in R 4.1).

`reg(formula_overall) |> fitstat()`

Hence, the `reg` function is called using the `formula_overall` as argument and 
the results of the `reg` model are are 'piped' to `fitstat`.

## formula

The key argument for `domin` is the formula input and understanding how it is 
parsed is important for the effective use of `domin`.

The formula argument is parsed using the `terms` utilities in the `stats` 
library.  Generally, 

and is not 'data frame aware'.  That is, shorthand such as `~ .` 
will not work to select variables in a data frame even if a `data` argument 
is supplied to the `domin` function.

```r{formula_example}
formula(mpg ~ .) |> terms(data=mtcars) |> formula()
```


...note all and sets go here as well

many modeling functions that accept a formula that follow the 
standard `response ~ terms` format. 

`domin` works closely with the `do.call` function and the structure of the 
`domin` function is similar as a result.  Users are encouraged to 
read the documentation for the `do.call` function to understand how `domin` is 
implemented and for bug/error checking.

wrapper function that can be used with 
The format used by domin can be extended 
to other functions focused on independent variable/predictor/feature-based RI 
can be accommodated with a additional wrapper functions based 
on the formula it creates and submits to modeling functions.  Some examples are
offered below.



The `{domir}` package is aimed at offering users tools for applying relative 
importance/RI analysis to a wide variety of statistical or machine learning 
functions.  In addition, this package aims to allow users to apply new 
fit statistics or metrics to statistical or machine learning 
functions that already have RI analysis implementations such as `lm`.

Currently, the only RI tool implemented in `{domir}` is a dominance analysis 
method `domin`. 
