---
title: Tools to Support Relative Importance Analysis
output: github_document
---



# Overview

The `domir` package provides a set of flexible wrapper and helper functions for 
conducting relative importance analysis with a focus on dominance analysis. 
The intention of this package is to provide tools that allow relative 
importance analysis across a wide variety of practical data analytic situations.

This documentation (for the time being) assumes some knowledge of the 
dominance analysis methodology and does not describe it extensively.

# Installation

To install the most recent stable version of `domir` from CRAN use:

`install.packages("domir")`

To install the working development version of `domir` using the `devtools` 
package use:

`devtools::install_github("https://github.com/jluchman/domir")`

# Details

The `domir` package is aimed at offering users tools for applying relative 
importance/RI analysis to a wide variety of statistical or machine learning 
functions.  In addition, this package aims to allow users to apply new 
fit statistics or metrics to statistical or machine learning 
functions that already have RI analysis implementations such as `lm`.

Currently, the only RI tool implemented in `domir` is a dominance analysis 
method `domin`. `domin` is a flexible wrapper function that can be used with 
many modeling functions that accept a formula that follow the 
standard `response ~ terms` format. The format used by domin can be extended 
to other functions focused on independent variable/predictor/feature-based RI 
can be accommodated with a additional wrapper functions based 
on the formula it creates and submits to modeling functions.  Some examples are
offered below.

# Examples

In this section, several examples of the `domin` function are outlined to 
illustrate how the function can accommodate a wide variety of different 
models--in some cases with a custom wrapper function.

## Linear Models with $R^2$

Fundamentally, `domir` is an extension of the `lmg` method in the 
`relaimpo` package and the example below is intended to illustrate how the 
differences between the packages by way of displayed information.

```{r replicate_relaimpo}
library(domir)

domin(mpg ~ am + vs + cyl, "lm", list("summary", "r.squared"), data=mtcars)

relaimpo::calc.relimp(lm(mpg ~ am + vs + cyl, data=mtcars))
```

## Linear Model with Alternative Fit Statistic

One, fairly basic, way that `domir` differs from other relative importance 
analysis modeling functions is in that it allows for different fit statistics 
other than the standard error variance $R^2$.  In this case, the McFadden 
pseudo-$R^2$ is used from the `pscl` package.

```{r lm_w_mcf}
domin(mpg ~ am + vs + cyl, "lm", list(function (model) 
    {pscl::pR2(model)},
        "McFadden"), data=mtcars)
```

## Ordered Logistic Regression

One analysis, to this point not supported in relative importance analysis 
software in R are models such as `polr` from the `MASS` package.

```{r da_polr}
mtcars2 <- data.frame(mtcars, carb2=as.factor(mtcars$carb))

domin(carb2 ~ am + vs + mpg, MASS::polr, 
    list(function(model) {pscl::pR2(model)}, "McFadden"), data=mtcars2)
```

## Multinomial Logistic (softmax) Regression

Another analysis, to this point not supported in relative importance analysis 
software in R are models such as `multinom` from the `nnet` package.

```{r da_multinom}
domin(carb2 ~ am + vs + mpg, nnet::multinom, 
    list(function(model) {pscl::pR2(model)}, "McFadden"), data=mtcars2)
```

## Decision Trees

`domir` can also accept models that have traditionally never been dominance 
analyzed such as `rpart` models.

Note that the function accepted by `domin`s `fitstat` argument can be 
quite complex as it is a wrapper to `do.call`.  Note that anonymous 
functions defined in the first element of `fitstat` must return a named 
list with a name that is callable in the second entry of `fitstat`.

```{r da_rpart}
mtcars2 <- data.frame(mtcars, carb2=as.factor(mtcars$carb))

domin(mpg ~ am + vs + cyl, rpart::rpart, 
    list(function(model) {list(R2=1-model$cptable[nrow(model$cptable),3])}, 
        "R2"), data=mtcars)
```

## Zero-Inflated Poisson

Complex models such as `zeroinfl` models from the package `pscl` can also 
be accommodated and as anonymous functions as well in `domin`s `reg` argument.

The below anonymous function in `reg` uses the entries in the formula to 
create a symmetric count and zero-inflation formulas.

The full model as submitted to `zeroinfl` looks like:

`art ~ fem + mar + kid5 + phd + ment | fem + mar + kid5 + phd + ment`

```{r da_zip}
domin(art ~ fem + mar + kid5 + phd + ment, 
        function(model, ...) {
            zip_terms <- paste(attr(terms(model), "term.labels"), collapse=" + ")
            zip_form <- paste(c(zip_terms, zip_terms), collapse=" | ")
            zip_form <- as.formula(paste(c(row.names(attr(terms(model), "factors"))[1], 
                zip_form), collapse=" ~ "))
            pscl::zeroinfl(zip_form, ...)
        },
    list(pscl::pR2, "McFadden"), data=pscl::bioChemists)

```

(more examples to come...)
